Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2018-10-06 analysis/paper/paper.Rmd
Replaces ShB with sheath blight in document

Sheath blight (ShB) disease (*Rhizoctonia solani* Kühn), anastomosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions he disease can cause up to 50\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009; @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen (N) management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

Because of the well documented effects of irrigation and N on ShB, we established field experiments to study what effects AWD irrigation technologies and N could have upon ShB incidence and severity in tropical rice.

@@

In both experiments, two sample areas, sized 1m x 1m, per plot were assessed. the total number of tillers per hill and number of tillers with ShB (incidence) were measured for nine hills per sample area. Tiller ShB severity was measured for four tillers per hill and six leaves tiller using a disease assessment rating scale for field evaluation developed at IRRI. The scale is an unequally spaced categorical scale where 0 - no disease; 1 - trace to 1% severity; 2 - 1 to 5%; 3 - 5 to 15%; 4 - 15 to 50%; 5 - 50 to 100%. Severity was considered to be the amount of leaf or sheath tissue covered by ShB lesions. Leaves were rated as green (living) having at least XXX% living tissue or dead, defined as having less than XXX% living tissue. The same rating methodology was used in both experiments. Five disease assessments were made in the 2015 experiment and four disease assessments were made in the 2016 experiment, respectively.

@@

The correlations between tiller sheath ShB incidence and tiller and leaf ShB severity were tested using Spearman’s correlation test, `cor.test`, [@R2018]. No correlation was found, so the analysis was completed for each variable independently with no assumed interaction.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller ShB incidence, 2015 tiller ShB severity, 2015 leaf ShB severity; 2016 tiller ShB incidence, 2016 tiller ShB severity; 2016 leaf ShB severity, respectively. The independent variables were the irrigation management and N rate with replicate treated as a random variable. The base levels (control treatments) for the analyses for 2015 were N rate - N0; irrigation management - FLD and for 2016, N rate - N60; irrigation management - FLD.

@@

In 2015 the incidence of tiller ShB remained low throughout the growing season (Fig. \@ref(fig:fig-1)a, \@ref(fig:fig-2)a). Water management was not significantly different (Fig. \@ref(fig:fig-3)a, \@ref(fig:fig-5)a). However, the N treatments, N100 and N120, were both significantly different when compared with the control N0 treatment (Fig. \@ref(fig:fig-4)a, \@ref(fig:fig-5)a).

@@

Tiller ShB severity remained below 2% (Fig. \@ref(fig:fig-2)c, \@ref(fig:fig-1)c). Both the N100 and N120 treatments were significantly different than the control N0 treatment (Fig. \@ref(fig:fig-4)c, \@ref(fig:fig-5)c). However, the AWD water management was not significantly different from the FLD treatment (Fig. \@ref(fig:fig-3)c, \@ref(fig:fig-5)c).

@@

Severity of leaf ShB remained low, less than 0.4% across all treatments (Fig. \@ref(fig:fig-1)e, \@ref(fig:fig-2)e). Both the N100 and N120 treatments were significantly different than the N0 treatment (Fig. \@ref(fig:fig-4)e, \@ref(fig:fig-5)e). However, the AWD water management was not significantly different from the FLD treatment (Fig. \@ref(fig:fig-3)e, \@ref(fig:fig-5)e).

@@

In 2016 the neither of the treatments, N rate or water management, were significantly different from the control treatments for leaf ShB severity (Fig. \@ref(fig:fig-3)f, \@ref(fig:fig-5)f, \@ref(fig:fig-4)f, \@ref(fig:fig-5)f).

@@

In both experiments we were unable to detect any significant effect of AWD on ShB that would cause increases in the disease, which could hinder adoption of the technology. In only one instance were we able to detect any effect of AWD on ShB. In the 2016 experiment the tiller ShB severity was significantly lower for AWD than FLD (Fig. \@ref(fig:fig-3)d, \@ref(fig:fig-5)d), indicating a possible adverse effect of using AWD on tiller ShB severity under conditions where there is high sheath blight pressure.

@@

The changes made for the 2016 experiment appear to have improved the experiment, increasing the incidence and decreased variability of sheath blight in the plots (Fig. \@ref(fig:fig-1)a:b, \@ref(fig:fig-2)c:d, \@ref(fig:fig-3)a:b, \@ref(fig:fig-4)c:d). The rice establishment method is known to affect the spread of sheath blight [@Willocquet2000]. It is possible that the mechanical transplanter could increase ShB spread in the canopy as compared with traditional manual transplanting. This however was not investigated as a part of this study but could bear further research.

>>>>

Sheath blight disease (*Rhizoctonia solani* Kühn), anastomosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions he disease can cause up to 50\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009; @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen (N) management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

Because of the well documented effects of irrigation and N on sheath blight, we established field experiments to study what effects AWD irrigation technologies and N could have upon sheath blight incidence and severity in tropical rice.

@@

In both experiments, two sample areas, sized 1m x 1m, per plot were assessed. the total number of tillers per hill and number of tillers with sheath blight (incidence) were measured for nine hills per sample area. Tiller sheath severity was measured for four tillers per hill and six leaves tiller using a disease assessment rating scale for field evaluation developed at IRRI. The scale is an unequally spaced categorical scale where 0 - no disease; 1 - trace to 1% severity; 2 - 1 to 5%; 3 - 5 to 15%; 4 - 15 to 50%; 5 - 50 to 100%. Severity was considered to be the amount of leaf or sheath tissue covered by sheath blight lesions. Leaves were rated as green (living) having at least XXX% living tissue or dead, defined as having less than XXX% living tissue. The same rating methodology was used in both experiments. Five disease assessments were made in the 2015 experiment and four disease assessments were made in the 2016 experiment, respectively.

@@

The correlations between tiller sheath sheath blight incidence and tiller and leaf sheath blight severity were tested using Spearman’s correlation test, `cor.test`, [@R2018]. No correlation was found, so the analysis was completed for each variable independently with no assumed interaction.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively. The independent variables were the irrigation management and N rate with replicate treated as a random variable. The base levels (control treatments) for the analyses for 2015 were N rate - N0; irrigation management - FLD and for 2016, N rate - N60; irrigation management - FLD.

@@

In 2015 the incidence of tiller sheath blight remained low throughout the growing season (Fig. \@ref(fig:fig-1)a, \@ref(fig:fig-2)a). Water management was not significantly different (Fig. \@ref(fig:fig-3)a, \@ref(fig:fig-5)a). However, the N treatments, N100 and N120, were both significantly different when compared with the control N0 treatment (Fig. \@ref(fig:fig-4)a, \@ref(fig:fig-5)a).

@@

Tiller sheath blight severity remained below 2% (Fig. \@ref(fig:fig-2)c, \@ref(fig:fig-1)c). Both the N100 and N120 treatments were significantly different than the control N0 treatment (Fig. \@ref(fig:fig-4)c, \@ref(fig:fig-5)c). However, the AWD water management was not significantly different from the FLD treatment (Fig. \@ref(fig:fig-3)c, \@ref(fig:fig-5)c).

@@

Severity of leaf sheath blight remained low, less than 0.4% across all treatments (Fig. \@ref(fig:fig-1)e, \@ref(fig:fig-2)e). Both the N100 and N120 treatments were significantly different than the N0 treatment (Fig. \@ref(fig:fig-4)e, \@ref(fig:fig-5)e). However, the AWD water management was not significantly different from the FLD treatment (Fig. \@ref(fig:fig-3)e, \@ref(fig:fig-5)e).

@@

In 2016 the neither of the treatments, N rate or water management, were significantly different from the control treatments for leaf sheath blight severity (Fig. \@ref(fig:fig-3)f, \@ref(fig:fig-5)f, \@ref(fig:fig-4)f, \@ref(fig:fig-5)f).

@@

In both experiments we were unable to detect any significant effect of AWD on sheath blight that would cause increases in the disease, which could hinder adoption of the technology. In only one instance were we able to detect any effect of AWD on sheath blight. In the 2016 experiment the tiller sheath blight severity was significantly lower for AWD than FLD (Fig. \@ref(fig:fig-3)d, \@ref(fig:fig-5)d), indicating a possible adverse effect of using AWD on tiller sheath blight severity under conditions where there is high sheath blight pressure.

@@

The changes made for the 2016 experiment appear to have improved the experiment, increasing the incidence and decreased variability of sheath blight in the plots (Fig. \@ref(fig:fig-1)a:b, \@ref(fig:fig-2)c:d, \@ref(fig:fig-3)a:b, \@ref(fig:fig-4)c:d). The rice establishment method is known to affect the spread of sheath blight [@Willocquet2000]. It is possible that the mechanical transplanter could increase sheath blight spread in the canopy as compared with traditional manual transplanting. This however was not investigated as a part of this study but could bear further research.
